---
title: "MTH6139 Time Series"
subtitle: "Coursework 1"
author:
- name: Hao Mei
date: "18 March 2024"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("QMlogo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:30%;')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(prophet)
library(zoo)
```

## Understanding CO2 Data

The dataset title "co2 '' put up info on the monthly concentration of
atmospheric carbon dioxide ( CO2 ) in parts per million , as record at
the Mauna Loa Observatory . These data play a crucial role in the
examination of climate change, since they provide insights into the
enduring patterns of rising carbon dioxide (CO2) concentrations in the
Earth's atmosphere.

## Purpose of the Project

The object lens of this study is to try and foreshadow the level of CO2
in order to savvy their pattern of increase and anticipate future
concentrations . Through this approach, we want to gain a deeper
understanding of the rate at which climate change is occurring and
assess the efficacy of strategies used to mitigate CO2 emissions.

Prepare the data for analysis with Prophet. Create a dataframe with two
columns: `ds` for the date and `y` for the CO2 levels.

```{r data-preparation}
co2.df = data.frame(
  ds = as.Date(as.yearmon(time(co2))), 
  y = co2)
```

## Model Building and Forecasting

With the data prepared, we proceed to fit a model and make forecasts.

```{r model-forecasting}
# Fit the model
m = prophet(co2.df)

# Generate future data frame for prediction
f = make_future_dataframe(m, periods=8, freq="quarter")

# Forecast
p = predict(m, f)

# Visualize the forecast
plot(m, p)
```

## Enhancing Our Analysis

Let's examine patterns and seasonality to better comprehend. Decomposing
the time series to see these components independently.

```{r decomposition}
prophet_plot_components(m, p)
```

```{r}
# Further Exploratory Data Analysis
# Plotting the raw data
plot(co2.df$ds, co2.df$y, type='l', main='Monthly CO2 Levels', xlab='Year', ylab='CO2 PPM', col='blue')

# Linear Trend Analysis
# A simple linear model to understand the growth of the series
linear_model <- lm(y ~ ds, data = co2.df)
summary(linear_model)

# Add linear trend line to the plot
abline(linear_model, col="red")
```
